Robotic cloth manipulation for clothing assistance task using Dynamic Movement Primitives

### 2017-02-10 paper/paper.tex
[major] add v3

% DOI
\acmDOI{To_be_inserted_after_paper_is_accepted}

@@

\begin{abstract}
	The need of robotic clothing assistance in the field of assistive robotics is growing, as it is one of the most basic activities in daily life of elderly and disabled people. In this study, we are investigating the applicability of using Dynamic Movement Primitives (DMP) as a task parameterization model for performing clothing assistance task. Robotic cloth manipulation task deals with putting a clothing article on both the arms. The clothing assistance is a challenging problem since robot must do cooperative manipulation by holding the cloth using both the arms while interacting with non-rigid and highly deformable clothes, and with the assisted person whose posture can vary during assistance. The system consists of dual arm Baxter research robot and Microsoft Kinect V2 RGBD sensor for tracking the posture of arms. The cloth manipulation is done by Baxter robot which follows trajectory generated by DMP system. We have performed experiments on soft mannequin instead of human. Result shows that DMPs are able to generalize movement trajectory for modified posture.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Due to the demographic trend in developed countries, robotic assistance in the field of elderly care in home environment is growing \cite{broekens2009assistive}. Although there has been a significant number of research done in this field, robotic clothing assistance is yet an open field for research. Not to mention that it is one of the basic and important assistance activities in daily life of elderly as well as disabled people. While rigid object manipulation with robots has mainly relied on precise robot control, deformable objects rather require complex control scheme. The clothing assistance is a challenging problem since robot is required to manage two difficulties: (a) robot must do cooperative manipulation by holding cloth using both the arms while interacting with non-rigid and highly deformable cloth and (b) maintain safe human-robot interaction with the assisted person whose posture can vary during assistance.

@@

\section{Related Works}
\label{sec:related_works}
There has been a significant number of research done in the field of robotic clothing assistance by using vision system. Klee et al. \cite{klee2015personalized} worked on personalized assistance for dressing a user. The robot request user to move towards robot, monitors motion using vision module and puts a hat on user once users is reachable by the robot. They haven't considered human and cloth interaction. Yamazaki et al. \cite{yamazaki2013method, yamazaki2014bottom}  worked on bottom dressing by a life-sized humanoid robot, where they recognize the cloth state by using optical flow on the images acquired from a single camera. They did not handle large occlusions. Yamakawa et al. \cite{yamakawa2011dynamic} proposed a new strategy for dynamic manipulation of sheet-like flexible objects by a high-speed robot system. The proposed system learns necessary motor skills from demonstration performed by a human subject. They worked on fast cloth state tracking but it is not related to robotic clothing assistance.

@@

The problem of robotic clothing assistance not only depends on tracking cloth dynamics but also motor skills required to perform the task. A combination of such techniques is encouraged. However, we see that there has been a large gap when compared it to the practical use cases in elderly care. The most important point should be learning rate of the task. We believe that \textit{Learning from Demonstration} frameworks are quick to learn in cases which involve complex task specific dynamics. The problem of putting sleeveless T-shirt into the arms of mannequin is close to the practical use case. Therefore, we are putting our efforts to solve it.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{flowchart_conf}
	\caption{Workflow of robotic cloth manipulation task. Initially, a demonstration is performed by moving Baxter arms in appropriate trajectory. This demonstration is recorded and parameterized by DMP. Later on, posture of the mannequin is changed and accordingly goal parameter of DMP is modified. Now, modified DMP can accommodate new posture.}
	\label{fig:workflow}
\end{figure*}

@@

\subsection{Robotic Cloth Manipulation using DMP}
\label{sec:system_overview}
In this section, we provide brief overview of our system. As per the formulation described in section \ref{sec:dmp}, DMP can learn from demonstration. Therefore we start by performing a demonstration by holding robot arm as shown in figure \ref{fig:workflow}. This is referred as ``Teaching Phase'', since in this phase, we are teaching skills to robot to perform the task.  During the demonstration, pose trajectory of end-effector is recorded using Baxter API and stored in a file. The term pose collectively refers to position and orientation. Once the demonstration is finished, DMP is initialized using recorded trajectory file. This is termed as ``Learn Trajectory'' phase. In this phase, recorded trajectory is parameterized by DMP system. The parameterized DMP can represent all the characteristics of original trajectory. Here, three DMP systems, one for each coordinate axis i.e., $x$, $y$ and $z$ are initialized for one arm. In this way, we have total six DMP systems, which can control both the arms of Baxter robot. In ``Testing Phase'', we change the posture of mannequin by lifting arms up or down. At this point, we use Kinect Sensor to get the 3D coordinates of arm. Before using Kinect Sensor, it is very important to do Kinect-Baxter calibration, so that 3D coordinates are translated from Kinect space to Baxter space. We change goal parameter of DMP trajectories by using this information. In this way, we have generalized DMP system, which can adapt modified posture referred as ``DMP Generalization''.

@@

\subsection{Clothing task using position DMP}
The aim of this experiment is to put sleeveless T-shirt on both the arms of mannequin by using DMP system. We use position data to initialize DMP trajectories, which are being used in this task. The posture of mannequin is changed by lifting arms up or down. At this point, we use Kinect Sensor to get 3D coordinates of arms. Now we change goal parameter of DMP trajectories by using this information. Modified DMP can be acquired by rolling out DMP system as described in section \ref{sec:dmp}.

In this experiment, the initialized DMP was modified to accommodate new posture by changing goal parameter of DMP. The generated trajectory was then run on Baxter robot as shown in figure \ref{fig:trajectory}. Newly generated DMP trajectory (shown in red color) was not only found well suited and capable of performing clothing task but also smooth compared to original trajectory (shown in blue color). A video demonstration of this experiment can be seen at YouTube\footnote{\url{http://youtu.be/Rb2JePazJjk}}.

@@

The clothing task has to deal with complex dynamics including manipulation of clothing article. Clothes are non-rigid, flexible and highly deformable objects, making the task more difficult to perform. During the task, we observed forces being applied at both the end-effector of Baxter robot. Trajectories were monitored and categorized into two \textit{success} and \textit{failure}. The mean of these two categories is calculated and plotted as shown in figure \ref{fig:position_force}. This is the average profile over several success and failure trajectories for different postures of the mannequin.

@@

\section{Discussion}
\label{sec:discussion}
The clothing assistance task deals with manipulation of highly deformable clothing article, which inherits complex dynamics. Failure can be detected by observing end-effector forces. The shape of clothing article keeps on changing during each trail, which affects the task settings. We noticed that there can be many failure scenarios such as clothing article gets stuck into the fingers, sleeve getting stuck on the arms, sleeve not entering the arm but entirely missing etc. The other reason of failure can be position based DMP system. DMP system was initialized by position data while ignoring the orientation of end-effector. The orientation was kept same as it was in original trajectory.

>>>>

% DOI
\acmDOI{to_be_inserted_after_paper_is_accepted}

@@

\begin{abstract}
	The need of robotic clothing assistance in the field of assistive robotics is growing, as it is one of the most basic and essential assistance activities in daily life of elderly and disabled people. In this study, we are investigating the applicability of using Dynamic Movement Primitives (DMP) as a task parameterization model for performing clothing assistance task. Robotic cloth manipulation task deals with putting a clothing article on both the arms. Robot trajectory varies significantly for various postures and also there can be various failure scenarios while doing cooperative manipulation with non-rigid and highly deformable clothing article. We have performed experiments on soft mannequin instead of human. Result shows that DMPs are able to generalize movement trajectory for modified posture.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Due to the demographic trend in developed countries, robotic assistance in the field of elderly care in home environment is growing \cite{broekens2009assistive}. Although there has been a significant number of research done in this field, robotic clothing assistance is yet an open field for research. It is one of the basic and essential assistance activities in daily life of elderly as well as disabled people. While rigid object manipulation with robots has mainly relied on precise robot control, deformable objects rather require complex control scheme. Clothing assistance is a challenging problem since robot is required to manage two difficulties: (a) robot must do cooperative manipulation by holding cloth using both the arms while interacting with non-rigid and highly deformable cloth and (b) maintain safe human-robot interaction with the assisted person whose posture can vary during assistance.

@@

\section{Related Works}
\label{sec:related_works}
There has been significant research done in the field of robotic clothing assistance by using vision system. Klee et al. \cite{klee2015personalized} worked on personalized assistance for dressing a user. The robot request user to move towards robot, monitors motion using vision module and puts a hat on user once users is reachable by the robot. They haven't considered human and cloth interaction. Yamazaki et al. \cite{yamazaki2013method, yamazaki2014bottom}  worked on bottom dressing by a life-sized humanoid robot, where they recognize the cloth state by using optical flow on the images acquired from a single camera. They did not handle large occlusions. Yamakawa et al. \cite{yamakawa2011dynamic} proposed a new strategy for dynamic manipulation of sheet-like flexible objects by a high-speed robot system. The proposed system learns necessary motor skills from demonstration performed by a human subject. They worked on fast cloth state tracking but it is not related to robotic clothing assistance.

@@

The problem of robotic clothing assistance not only depends on tracking cloth dynamics but also motor skills required to perform the task. A combination of such techniques is encouraged. However, we see that there has been a large gap when compared to the practical use cases in elderly care. The most important point should be learning rate of the task. We believe that \textit{Learning from Demonstration} frameworks are quick to learn in cases which involve complex task specific dynamics. The problem of putting sleeveless T-shirt into the arms of mannequin is close to the practical use case. Therefore, we are putting our efforts to solve it.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{flowchart_conf}
	\caption{Workflow of robotic cloth manipulation task. Initially, a kinesthetic demonstration is performed with the robot controlled in gravity compensation mode. This demonstration is recorded and parameterized by DMP. Later on, posture of the mannequin is changed and accordingly start and goal parameter of DMP is modified. Now, modified DMP can accommodate new posture.}
	\label{fig:workflow}
\end{figure*}

@@

\subsection{Robotic cloth manipulation using DMP}
\label{sec:system_overview}
In this section, we provide brief overview of our system. As per the formulation described in section \ref{sec:dmp}, DMP can learn from demonstration. Therefore we start by performing a kinesthetic demonstration with the robot controlled in gravity compensation mode as shown in figure \ref{fig:workflow}. This is referred as ``Teaching Phase'', since in this phase, we are teaching skills to robot to perform the task.  During the demonstration, pose trajectory of end-effector is recorded using Baxter API and stored in a file. The term pose collectively refers to position as cartesian position $p = (p_x, p_y, p_z) \in \mathbb{R}^3$ and orientation as quaternion $q = (q_x, q_y, q_z, q_w) \in \mathbb{R}^4$. Once the demonstration is finished, DMP is parameterized using recorded trajectory file. This is termed as ``Learn Trajectory'' phase. The parameterized DMP can represent all the characteristics of original trajectory. Here, three DMP systems, one for each coordinate axis i.e., $x$, $y$ and $z$ are initialized for one arm. In this way, we have total six DMP systems, which can control both the arms of Baxter robot. In ``Testing Phase'', we change the posture of mannequin by changing the angle of inclination as shown in figure \ref{fig:accuracy}. At this point, we use Kinect Sensor to get the 3D coordinates of wrist and shoulder of mannequin. We change start and goal parameter of DMP trajectories by using this information. In this way, we have generalized DMP system, which can adapt modified posture referred as ``DMP Generalization''. Before using Kinect Sensor, it is extremely important to do Kinect-Baxter calibration, so that 3D coordinates are translated from Kinect space to Baxter space. For Kinect-Baxter calibration, we collect a dataset of points observed by Baxter and Kinect. Then we use absolute orientation calibration \citep{umeyama1991least} to align the frames of reference.

We have divided the complete trajectory into two parts: (a) The reaching part, which refers to the trajectory starts from home position of robot and ends till hands of mannequin (b) the clothing part, which refers to the trajectory starts from hands of mannequin and reaches upto shoulder nearly. The reaching part can be performed through simple position based controller but for the clothing phase, we need to train DMP since this part changes drastically between postures and can be prone to failure.

@@

\subsection{Clothing task using position DMP}
The aim of this experiment is to put sleeveless T-shirt on both the arms of mannequin by using DMP system. We use position data to initialize DMP trajectories, which are being used in this task. The posture of mannequin is changed. At this point, we use Kinect Sensor to get 3D coordinates of wrist and shoulder of mannequin. Now we change start and goal parameter of DMP trajectories by using this information. Modified DMP can be acquired by rolling out DMP system as described in section \ref{sec:dmp}.

In this experiment, the initialized DMP was modified to accommodate new posture by changing start and goal parameter of DMP. The generated trajectory was then run on Baxter robot as shown in figure \ref{fig:trajectory}. Newly generated DMP trajectory (shown in red color) was not only found well suited and capable of performing clothing task but also smooth compared to original trajectory (shown in blue color). A video demonstration of this experiment can be seen at YouTube\footnote{\url{http://youtu.be/Rb2JePazJjk}}.

@@

The clothing task has to deal with complex dynamics including manipulation of clothing article. Clothes are non-rigid, flexible and highly deformable objects, making the task more difficult to perform. During the task, we observed forces being applied at both the end-effector of Baxter robot. Trajectories were monitored and categorized into two \textit{success} and \textit{failure}. The mean of these two categories is calculated and plotted as shown in figure \ref{fig:position_force}. Force value is the norm of force applied in all three cartesian directions. This is the average profile over several \textit{success} and \textit{failure} trajectories for different postures of left arm of mannequin.

@@

\section{Discussion}
\label{sec:discussion}
The clothing assistance task deals with manipulation of highly deformable clothing article, which inherits complex dynamics. Failure can be detected by observing end-effector forces. The shape of clothing article keeps on changing during each trail, which affects the task settings. That is why, a trajectory which was able to preform successfully in the past, doesn't work later sometimes. We noticed that there can be many failure scenarios such as clothing article gets stuck into the fingers, sleeve getting stuck on the arms, sleeve not entering the arm but entirely missing etc. Proposed failure detection method by using force information can detect failures when clothing article is stuck tightly. The other reason of failure can be position based DMP system. Orientation plays a major role in the task however DMP system was initialized by position data while ignoring the orientation of end-effector. The orientation was kept same as it was in original trajectory.
